Data leakage - does imputation before or after splitting matter?

### 2022-10-12 workflow/notebooks/data-leakage-imputation.Rmd

TODO histogram showing # of NA's in OTU columns instead of barplot (too many OTUs for barplot)
```{r plot_histogram}
missing.values <- allcause_full_dat %>%
    gather(key = "key", value = "val") %>%
    mutate(is.missing = is.na(val)) %>%
    group_by(key, is.missing) %>%
    summarise(num.missing = n()) %>%
    filter(is.missing==T) %>%
    select(-is.missing) %>%
    arrange(desc(num.missing)) 
missing.values %>%
  ggplot() +
    geom_bar(aes(x=key, y=num.missing), stat = 'identity') +
    labs(x='variable', y="number of missing values", title='Number of missing values') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r check_works_with_simple_frame}
dat <- data.frame(x1 = c(NA, 5, 5, NA, 1, 2),
                   x2 = c(1, 2, 3, NA, 5, 6),
                   x3 = 1)
temp <- colSums(is.na(dat)) 
head(temp)
sum(temp)
missing.values <- dat %>%
    gather(key = "key", value = "val") %>%
    mutate(is.missing = is.na(val)) %>%
    group_by(key, is.missing) %>%
    summarise(num.missing = n()) %>%
    filter(is.missing==T) %>%
    select(-is.missing) %>%
    arrange(desc(num.missing)) 
missing.values %>%
  ggplot() +
    geom_bar(aes(x=key, y=num.missing), stat = 'identity') +
    labs(x='variable', y="number of missing values", title='Number of missing values') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

>>>>

TODO histogram showing # of NA's in OTU columns instead of barplot (too many OTUs for barplot)

TODO functions to find NAs and create histogram

```{r plot_histogram}
missing.values <- allcause_full_dat %>%
    gather(key = "key", value = "val") %>%
    mutate(is.missing = is.na(val)) %>%
    group_by(key, is.missing) %>%
    summarise(num.missing = n()) %>%
    filter(is.missing==T) %>%
    select(-is.missing) %>%
    arrange(desc(num.missing)) 
missing.values %>%
  ggplot() +
    geom_bar(aes(x=key, y=num.missing), stat = 'identity') +
    labs(x='variable', y="number of missing values", title='Number of missing values') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r check_works_with_simple_frame}
dat <- data.frame(x1 = c(NA, 5, 5, NA, 1, 2),
                   x2 = c(1, 2, 3, NA, 5, 6),
                   x3 = 1)
temp <- colSums(is.na(dat)) 
head(temp)
sum(temp)
missing.values <- dat %>%
    gather(key = "key", value = "val") %>%
    mutate(is.missing = is.na(val)) %>%
    group_by(key, is.missing) %>%
    summarise(num.missing = n()) %>%
    filter(is.missing==T) %>%
    select(-is.missing) %>%
    arrange(desc(num.missing)) 
missing.values %>%
  ggplot() +
    geom_bar(aes(x=key, y=num.missing), stat = 'identity') +
    labs(x='variable', y="number of missing values", title='Number of missing values') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
